On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-27 TTau_Halpha/kurosawa.tex
Minior update
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@640 25464f21-6099-402c-b30a-640f1d40d38b

    \includegraphics[%
      clip,
      scale=0.45,angle=270]{figures/discwind_hybrid_ha_image.eps}

@@

\citet{reipurth:1996} found that Type I profiles (symmetric around line centres) constituted 26~per~cent
of the 43 CTTS H$\alpha$ profiles in their sample, making it the second most common
morphological type. We find that Type~I  profiles are dominated by magnetospheric
emission at rates in excess of $\appox 10^{-8}\, \dot{M_\odot}\,{\rm yr}^{-1}$. Type~I profiles
appear at a wide range of inclinations (Figure~\ref{magnetosphere only inc dependence figure}).
An object that displays a Type~I profile may have relatively weak disc wind, but in this
case it must be viewed from the pole (in order that no wind material is seen projected
against the stellar photosphere).

@@

The right panel in Figure~\ref{fig:effect_inclination_compare} shows
the profiles computed with the disc-wind, magnetosphere hybrid model
(section~\ref{sub:result-discwind-knigge}) 
at inclination angles $i=10^{\circ}$, $55^{\circ}$ and
$80^{\circ}$. The magnetospheric accretion component uses the  
parameters used as the reference model (Table~\ref{tab:std_parameters}).  
The figure shows that the absorption
feature becomes stronger as the inclination increases, mainly because
of the geometrical configuration.  The optical depth to becomes larger
as the inclination angle increases since there is more material
towards the equatorial plane.  Although not shown here, this tendency
of the inclination dependency holds for the models with the wide
ranges of $\beta$ and $T_{\mathrm{wind}}$ values.  
As a result, the EW becomes smaller as the inclination
angle increases. Clearly, this is not consist with the trend seen in
the observation of \citet{appenzeller:2005}. Recall that a similar
inclination dependency was found for the models with only
magnetosphere (Figure~\ref{fig:acc_inclination_effect}).

@@

The right panel in Figure~\ref{fig:effect_inclination_compare} shows
the H$\alpha$ profiles computed using the alternative hybrid model at
the same inclination angles used for the disc-wind-magnetosphere
hybrid model. The wind acceleration parameter ($\beta$) and the wind
temperature ($T_{\mathrm{wind}}$)used here are $0.5$ and
$9000\,\Kelvin$ respectively. All the other parameters are same as in
the disc-wind magnetosphere model.  With this set of $\beta$ and
$T_{\mathrm{wind}}$, the wind emission is comparable or a slightly
greater than the emission from the magnetosphere. The profiles show
rather complex absorption features because of the interaction of
photons created in the wind with the magnetosphere. 
While the blue-shifted absorption features are formed in the wind, the
absorption features near the line centre and in the red wing (only for
$i=10^\circ$) are caused by the wind emission by the magnetosphere
which has a rather complex flow geometry. 
Because of the geometry of the wind, the P-Cyg absorption feature
weakens as the inclination increases. The optical depth of the wind is
much higher in 
the polar direction compared to that at higher inclination. For the
same reason, the EW of the line increases as the inclination increase
as clearly seen in the figure.  Although not shown here, this tendency
of the inclination dependency holds for the models with this wind
models for wide ranges of $\beta$ and $T_{\mathrm{wind}}$ as long as
the emission from the wind is comparable to that of the magnetosphere.

@@

The profiles computed with the hybrid model showed that the wind
emission contribution becomes significantly larger than that from the
magnetosphere (Figure~\ref{atlas_hybrid} and
Table~\ref{tab:ew_ratio}).  Henceforth, a large H$\alpha$ equivalent
width does not necessary indicate a large mass-accretion rate, but it
could be caused by the mass-loss by the wind outflow.

@@

The value of $\mu$, 0.1,  suggested by both MHD calculations and observations
(c.f.~\citep{koenigl:2000}) is consistent with the 
disc-wind-magnetosphere models presented here. We found that the
hybrid models with $0.05<\mu<0.2$ are also consistent with the type
of the profiles seen in the observations (e.g~\citealt{reipurth:1996}).

@@

>>>>

    \includegraphics[%
      clip,
      scale=0.45,angle=270]{figures/discwind_hybrid_ha_image2.eps}

@@

\citet{reipurth:1996} found that Type I profiles (symmetric around
line centres) constituted 26~per~cent of the 43 CTTS H$\alpha$
profiles in their sample, making it the second most common
morphological type. We find that Type~I profiles are dominated by
magnetospheric emission at rates in excess of $\approx 10^{-8}\,
\dot{M_\odot}\,\mathrm{yr^{-1}}$. Type~I profiles appear at a wide range
of inclinations (Figure~\ref{fig:acc_inclination_effect}).  An object
that displays a Type~I profile may have 
relatively weak disc wind, but in this case it must be viewed from the
pole (in order that no wind material is seen projected against the
stellar photosphere).

@@

The right panel in Figure~\ref{fig:effect_inclination_compare} shows
the profiles computed with the disc-wind, magnetosphere hybrid model
(section~\ref{sub:result-discwind-knigge}) 
at inclination angles $i=10^{\circ}$, $55^{\circ}$ and
$80^{\circ}$. The magnetospheric accretion component uses the  
parameters used as the reference model (Table~\ref{tab:std_parameters}).  
The figure shows that the absorption
feature becomes stronger as the inclination increases, mainly because
of the geometrical configuration.  The optical depth to becomes larger
as the inclination angle increases since there is more material
towards the equatorial plane.  This tendency
of the inclination dependency holds for the models with the wide
ranges of $\beta$ and $T_{\mathrm{wind}}$ values.  
As a result, the EW becomes smaller as the inclination
angle increases: 41\AA, 27\AA, and 19\AA\, for $i=10^\circ$, $55^\circ$, and
$80^\circ$ respectively. 
Clearly, this is not consist with the trend seen in
the observation of \citet{appenzeller:2005}. Recall that a similar
inclination dependency was found for the models with only
magnetosphere (Figure~\ref{fig:acc_inclination_effect}).

@@

The right panel in Figure~\ref{fig:effect_inclination_compare} shows
the H$\alpha$ profiles computed using the alternative hybrid model at
the same inclination angles used for the disc-wind-magnetosphere
hybrid model. The wind acceleration parameter ($\beta$) and the wind
temperature ($T_{\mathrm{wind}}$)used here are $0.5$ and
$9000\,\Kelvin$ respectively. All the other parameters are same as in
the disc-wind magnetosphere model.  With this set of $\beta$ and
$T_{\mathrm{wind}}$, the wind emission is comparable or a slightly
greater than the emission from the magnetosphere. The profiles show
rather complex absorption features because of the interaction of
photons created in the wind with the magnetosphere. 
While the blue-shifted absorption features are formed in the wind, the
absorption features near the line centre and in the red wing (only for
$i=10^\circ$) are caused by the wind emission by the magnetosphere
which has a rather complex flow geometry. 
Because of the geometry of the wind, the P-Cyg absorption feature
weakens as the inclination increases. The optical depth of the wind is
much higher in 
the polar direction compared to that at higher inclination. For the
same reason, the EW of the line increases as the inclination increase
as clearly seen in the figure. Specifically their values are 
49\AA, 60\AA, and 84\AA\, for $i=10^\circ$, $55^\circ$, and
$80^\circ$ respectively.  This tendency
of the inclination dependency holds for the models with this wind
models for wide ranges of $\beta$ and $T_{\mathrm{wind}}$ as long as
the emission from the wind is comparable to that of the magnetosphere.

@@

The profiles computed with the hybrid model showed that the wind
emission contribution becomes significantly larger than that from the
magnetosphere (Figure~\ref{fig:atlas_hybrid} and
Table~\ref{tab:ew_ratio}).  Henceforth, a large H$\alpha$ equivalent
width does not necessary indicate a large mass-accretion rate, but it
could be caused by the mass-loss by the wind outflow.

@@

The value of $\mu$, 0.1,  suggested by both MHD calculations and observations
(c.f.~\citealt{koenigl:2000}) is consistent with the 
disc-wind-magnetosphere models presented here. We found that the
hybrid models with $0.05<\mu<0.2$ are also consistent with the type
of the profiles seen in the observations (e.g~\citealt{reipurth:1996}).

@@

%
% Appendix starts here
%
%\newpage
%\onecolumn
\appendix

\section{H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model}

Here we present xxxxxxxxxxxxxxx.

%\twocolumn
%======================================================================
\begin{figure}

  \begin{center}
      \includegraphics[%
	clip,
	scale=0.6]{figures/discwind_hybrid_mdot7_inc10.eps}\tabularnewline
      \vspace{0.5 cm}

      \includegraphics[%
	clip,
      scale=0.6]{figures/discwind_hybrid_mdot7_inc55.eps}\tabularnewline

      \vspace{0.5 cm}

      \includegraphics[%
	clip, scale=0.6]{figures/discwind_hybrid_mdot7_inc80.eps}\tabularnewline
  \end{center}

  \caption{The H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model with
  $\dot{M_{\mathrm{acc}}} =10^{-7}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-6}\,M_{\sun}\,\mathrm{yr{-1}}$. }

  \label{fig:atlas01}

\end{figure}
%======================================================================

%======================================================================
\begin{figure}

  \begin{center}
      \includegraphics[%
	clip,
	scale=0.6]{figures/discwind_hybrid_mdot8_inc10.eps}\tabularnewline
      \vspace{0.5 cm}

      \includegraphics[%
	clip,
      scale=0.6]{figures/discwind_hybrid_mdot8_inc55.eps}\tabularnewline

      \vspace{0.5 cm}

      \includegraphics[%
	clip, scale=0.6]{figures/discwind_hybrid_mdot8_inc80.eps}\tabularnewline
  \end{center}

  \caption{The H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model with
  $\dot{M_{\mathrm{acc}}} =10^{-8}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-7}\,M_{\sun}\,\mathrm{yr{-1}}$. }

  \label{fig:atlas02}

\end{figure}
%======================================================================

%======================================================================
\begin{figure}

  \begin{center}
      \includegraphics[%
	clip,
	scale=0.6]{figures/discwind_hybrid_mdot9_inc10.eps}\tabularnewline
      \vspace{0.5 cm}

      \includegraphics[%
	clip,
      scale=0.6]{figures/discwind_hybrid_mdot9_inc55.eps}\tabularnewline

      \vspace{0.5 cm}

      \includegraphics[%
	clip, scale=0.6]{figures/discwind_hybrid_mdot9_inc80.eps}\tabularnewline
  \end{center}

  \caption{The H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model with
  $\dot{M_{\mathrm{acc}}} =10^{-9}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-10}\,M_{\sun}\,\mathrm{yr{-1}}$. }

  \label{fig:atlas03}

\end{figure}
%======================================================================
